---
title: "simulation"
author: "Asif Enan"
format: html
editor: visual
---

## Using simulation to answer some probability questions

```{r}
library(tidyverse)
source("craps-code.R")
```

### 1) What happens to the probability of winning for increasing number of games?

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
# Set up simulation counts
# simulation_sizes <- seq(100,5000,100) this was taking a lot of time so I am gonna use a shorter version of this
simulation_sizes <- seq(100,2000,50)

# Running simulations and calculate winning probabilities for each simulations_size
win_probabilities <- map_dfr(simulation_sizes, ~ {
  # Simulating N games
  simulation_result <- run_craps_simulation(.x)
  
  # Calculating win probability
  win_prob <- mean(simulation_result$Match_Outcome == "Win")
  
  # Returning the results as a tibble
  tibble(Simulation_Size = .x, Win_Probability = win_prob)
})


print(win_probabilities)


```

Graphing the results from above

```{r}
win_probabilities %>%
  ggplot(aes(x=Simulation_Size,y=Win_Probability))+
  geom_line()
```

The probability of winning seems to be stabilizing \[i.e. variations getting less\] as we increase the simulation size

### 2) If we play 20 thousand game of craps, which of the 4 outcomes occurs the most?

```{r}
simulation_result20k <- run_craps_simulation(20000)

# Categorizing game outcomes
outcome_breakdown <- simulation_result20k %>%
  mutate(Outcome_Category = case_when(
    is.na(Point_Value) & Match_Outcome == "Win" ~ "Natural Win",
    is.na(Point_Value) & Match_Outcome == "Loss" ~ "Craps Loss",
    !is.na(Point_Value) & Match_Outcome == "Win" ~ "Point Hit Win",
    !is.na(Point_Value) & Match_Outcome == "Loss" ~ "Seven-Out Loss"
  )) %>%
  group_by(Outcome_Category) %>%
  summarize(Count = n(),
            Probability = Count / nrow(simulation_result))

# View results
print(outcome_breakdown)

```

We see that the \`Seven-out Loss\` occurs the most having the highest probability.

### 3) If we play 20 thousand game of craps, what is the average number of rolls per game?

```{r}
average_rolls <- simulation_result20k %>%
  summarize(Average_Number_of_Rolls = mean(Number_of_Rolls))

# View result
print(average_rolls)

```

Average number of rolls per game is 3.367 which means by that time either Point-hit Win or Seven-Out Loss
